---
title: "Bond Future Data HAR Model"
author: "Shawn Xiao"
date: "2025-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(zoo)
library(dplyr)
library(visdat)
library(tidyr)
library(quantmod)
library(tidyquant)
library(ggplot2)
```


#Pulling Data and some data processing
```{r pressure, echo=FALSE}

# List of US Bond Futures (Symbol -> Description)

# Yahoo Finance bond futures symbols and descriptions
bond_futures_df <- data.frame(
  Symbol = c("ZT=F", "Z3N=F", "ZF=F", "ZN=F", "TN=F", "TWE=F", "ZB=F", "UB=F"),
  Description = c(
    "2-Year Treasury Note Futures: Notes with 1y9m–2y maturity",
    "3-Year Treasury Note Futures: Notes with 2y9m–3y maturity",
    "5-Year Treasury Note Futures: Notes with 4y2m–5y3m maturity",
    "10-Year Treasury Note Futures: Notes with 6y6m–8y maturity",
    "Ultra 10-Year Treasury Note Futures: Notes with 9y5m–10y maturity",
    "20-Year Treasury Bond Futures: Bonds with 19y2m–19y11m maturity",
    "30-Year Treasury Bond Futures: Bonds with 15y+ maturity",
    "Ultra 30-Year Treasury Bond Futures: Bonds with 25y+ maturity"
  ),
  row.names = c("ZT=F", "Z3N=F", "ZF=F", "ZN=F", "TN=F", "TWE=F", "ZB=F", "UB=F")
)

# View the data frame
print(bond_futures_df)

# Get bond futures
bond_futures <- tq_get("ZB=F", from = "2005-01-01", to = Sys.Date())

# Plot it
ggplot(bond_futures, aes(x = date, y = adjusted)) +
  geom_line() +
  labs(title = "30-Year T-Bond Futures", x = "Date", y = "Adjusted Price")

vis_miss(bond_futures)
missing_values <- is.na(bond_futures) 

#Missing Data due to US.Market Non-trading Day, best to carry on the previous day data

bond_futures = bond_futures %>% fill(adjusted)
  sum(is.na(bond_futures$adjusted))


```

#HAR Model Pesudo Code
```{r}
#Log Return
bond_futures <- bond_futures %>%
  mutate(
    log_return = log(adjusted / lag(adjusted)),
    RV = log_return
  )

#Since Log return, first value is output as 0)
sum(is.na(bond_futures$log_return))
length(bond_futures$adjusted)

#Create HAR lag
har_df <- bond_futures %>%
  mutate(
    RV_D = lag(RV, 1),
    RV_W = rollmean(lag(RV, 1), k = 5, fill = NA, align = "right"),
    RV_M = rollmean(lag(RV, 1), k = 22, fill = NA, align = "right")
  ) %>%
  drop_na()

#Fit HAR model
har_model <- lm(RV ~ RV_D + RV_W + RV_M, data = har_df)

#Summary of model
summary(har_model)

#Add Prediction, since we are predicting, one month of data is being used to forecast at very start, so the dataframe start around Feb
har_df <- har_df %>%
  mutate(RV_hat = predict(har_model))

ggplot(har_df, aes(x = date)) +
  geom_line(aes(y = RV), color = "black", alpha = 0.6) +
  geom_line(aes(y = RV_hat), color = "blue") +
  labs(
    title = "HAR Model: Actual vs Predicted Realized Variance",
    y = "Squared Log Return (RV)", x = "Date"
  )
```

```{r}
#Log Return
bond_futures <- bond_futures %>%
  mutate(
    log_return = log(adjusted / lag(adjusted)),
    # Lacking Intrday Data, Wanna make a comparison of betweeen Orginial and daily proxy for realized variance using log return squared
    RV = log_return^2 
  )

#Since Log return, first value is output as 0)
sum(is.na(bond_futures$log_return))
length(bond_futures$adjusted)

#Create HAR lag
har_df <- bond_futures %>%
  mutate(
    RV_D = lag(RV, 1),
    RV_W = rollmean(lag(RV, 1), k = 5, fill = NA, align = "right"),
    RV_M = rollmean(lag(RV, 1), k = 22, fill = NA, align = "right")
  ) %>%
  drop_na()

#Fit HAR model
har_model <- lm(RV ~ RV_D + RV_W + RV_M, data = har_df)

#Summary of model
summary(har_model)

#Add Prediction, since we are predicting, one month of data is being used to forecast at very start, so the dataframe start around Feb
har_df <- har_df %>%
  mutate(RV_hat = predict(har_model))

ggplot(har_df, aes(x = date)) +
  geom_line(aes(y = RV), color = "black", alpha = 0.6) +
  geom_line(aes(y = RV_hat), color = "blue") +
  labs(
    title = "HAR Model: Actual vs Predicted Realized Variance",
    y = "Squared Log Return (RV)", x = "Date"
  )
```


Error in `$<-`:
! Assigned data `predict(har_model)` must be compatible with existing data.
✖ Existing data has 5124 rows.
✖ Assigned data has 5073 rows.
ℹ Only vectors of size 1 are recycled.
Caused by error in `vectbl_recycle_rhs_rows()`:
! Can't recycle input of size 5073 to size 5124.
Run `rlang::last_trace()` to see where the error occurred.
Error in `$<-`(`*tmp*`, RV_hat, value = c(`1` = 3.16596415191453e-05,  : 
  
✖ Existing data has 5124 rows.
✖ Assigned data has 5073 rows.
ℹ Only vectors of size 1 are recycled.
Caused by error in `vectbl_recycle_rhs_rows()`:
! Can't recycle input of size 5073 to size 5124.
